### 2019-02-27 sections/intuition.tex
fit in more

Input events drive evaluation of tasks.
When the system receives a valid event, it applies the event to the current task, which results in a new task.
In this way the system communicates with the environment.
Input events are synchronous.
We assume systems react instantaneous to each such event.

@@

The cigarette smokers problem \cite{books/Downey08LBOS} is a surprisingly tricky synchronisation problem.
We study it here because it demonstrates the capabilities of guarded steps.
% The problem requires waiting for two conditions, waking up only if both conditions are satisfied.
The problem is stated as follows.
In order to smoke a cigarette, three ingredients are required: tobacco, paper, and a match.
There are three smokers, each having one of the ingredients and requiring the other two.
There is an agent that randomly provides two of the ingredients.
The difficulty lies in the requirement that only the smoker may proceed whose missing ingredients are present.

Downey models availability of the ingredients with a semaphore for each ingredient.
The agent randomly signals two of the three semaphores.
The solution proposed by Downey involves an additional mutex, three additional semaphores, three additional threads called \emph{pushers}, and three regular Boolean variables.
The job of the pushers is to record availability of their ingredient in their Boolean variable, and check availability of other resources, waking the correct smoker when appropriate.

@@

\TOPHAT allows a simple solution to this problem, using guarded steps.
Steps can be guarded with arbitrary expressions, and the parallel combinator can be used to watch two shared editors at the same time.
Let \TS{match}, \TS{paper}, and \TS{tobacco} be references to Booleans.
The smokers are defined as follows.
\begin{TASK}
  let continue = \<<x,y>>. if x /\ y then smoke else fail in
  let tobaccoSmoker = (update match <&> update paper) >>? continue in
  let paperSmoker = (update tobacco <&> update match) >>? continue in
  let matchSmoker = (update tobacco <&> update paper) >>? continue
\end{TASK}
When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled.

>>>>

Input events drive evaluation of tasks.
When the system receives a valid event, it applies this event to the current task, which results in a new task.
In this way the system communicates with the environment.
Inputs are synchronous.
We assume systems react instantaneous to each such event.

@@

The cigarette smokers problem \cite{books/Downey08LBOS} is a surprisingly tricky synchronisation problem.
We study it here because it demonstrates the capabilities of guarded steps.
% The problem requires waiting for two conditions, waking up only if both conditions are satisfied.
The problem is stated as follows.
In order to smoke a cigarette, three ingredients are required: tobacco, paper, and a match.
There are three smokers, each having one of the ingredients and requiring the other two.
There is an agent that randomly provides two of those.
The difficulty lies in the requirement that only the smoker may proceed whose missing ingredients are present.

Downey models availability of the ingredients with a semaphore for each ingredient.
The agent randomly signals two of the three.
The solution proposed by Downey involves an additional mutex, three additional semaphores, three additional threads called \emph{pushers}, and three regular Boolean variables.
The job of the pushers is to record availability of their ingredient in their Boolean variable, and check availability of other resources, waking the correct smoker when appropriate.

@@

\TOPHAT allows a simple solution to this problem, using guarded steps.
Steps can be guarded with arbitrary expressions.
The parallel combinator can be used to watch two shared editors at the same time.
Let \TS{match}, \TS{paper}, and \TS{tobacco} be references to Booleans.
The smokers are defined as follows.
% The problem is then specified as follows.
\begin{TASK}
  let continue = \<<x,y>>. if x /\ y then smoke else fail in
  let tobaccoSmoker = (update match <&> update paper) >>? continue in
  let paperSmoker = (update tobacco <&> update match) >>? continue in
  let matchSmoker = (update tobacco <&> update paper) >>? continue
\end{TASK}
When the agent supplies two of the ingredients by setting the respective shares to \TS{True}, only the step of the smoker that waits for those becomes enabled.

### 2019-02-27 sections/language.tex
fit in more

Consider the following program.
\begin{TASK}
  enter Int >>= \n. if n == 42 then edit "Good" else edit "Bad"
\end{TASK}
Initially the step is guarded because the editor doesn't have a value.
When the user enters an integer, the program continues immediately with either \TS|edit "Good"| or \TS|edit "Bad"|, depending on the input.

@@

External choice ($\Xor$) is different in this regard.
An external choice requires users to pick a branch before continuing with it.
This means users cannot work on the branches of an external choice before picking one. \todo{Alejandro: Is this actually implied?}

>>>>

Consider the following program.
\begin{TASK}
  enter Int >>= \n. if n == 42 then edit "Good" else edit "Bad"
\end{TASK}
Initially, the step is guarded because the editor does not have a value.
When users enter an integer, the program continues immediately with either \TS{edit "Good"} or \TS{edit "Bad"}, depending on the input.

@@

External choice ($\Xor$) is different in this regard.
An external choice requires users to pick a branch before continuing with it.
This means users cannot work on the branches of an external choice before picking one.
%XXX: Alejandro: Is this actually implied?
% Ik snap niet zo goed wat Alejandro hier mee bedoelt... --TS
